---
title: "Homework 4"
author: "Jiawen Qi (jiq10)"
date: "November 1, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Read in the iris data

In RStudio, there is iris data.

```{r, warning=FALSE}
iris <- iris
summary(iris) # give a summary to the iris
```

## Step 2: Randomly pick 50 observations to be the testing dataset

```{r, warning=FALSE}
testing <- iris[sample(1:150, 50),] # get a random testing dataset
```

## Step 3: Do the classification method 100 times

```{r, warning=FALSE}
library(class) # open the class to use knn
for (i in 1:100) { # do this loop 100 times
  training <- iris[sample(1:150, 100, replace = TRUE),] # taking a random sample of a size 100 from iris as training with replacement
  # build a model predicting what type of iris the observation is given its sepal width
  train.data <- as.matrix(training$Sepal.Width) # get the value of sepal width
  train.class <- as.matrix(training$Species) # get the class
  test.data <- as.matrix(testing$Sepal.Width) # get the date to be predicted
  knn5 = knn(train.data,            # training data as a matrix
             test.data,             # testing data as a matrix
             train.class,           # class for training data
             k=5)                   # k = 5
  test.predicted <- data.frame(test.data, pred = knn5) # get predicted values for test
  testing[,i+5] <- test.predicted$pred # merge predicted into iris
}
```

## Step 4: Count the predict results

I will count the frequency of predict results for each obsercation. And treat the class with highest frequency as the predicted result.

```{r, warning=FALSE}
results <- testing[,c(5:105)]
results <- cbind.data.frame(predictAs = 0, results) # create  columns to count the results
row.names(results) <- c(1:50) # rename the row
for (i in 1:50) {
  predictCount<- table(t(results[i,3:102])) # get the frequency
  predictCountFrame <- as.data.frame(predictCount) # change to a data frame
  predictClass <- predictCountFrame[predictCountFrame$Freq == max(predictCountFrame$Freq),1] # get the class of the high frequency
  predictClass <- as.character(predictClass)
  results$predictAs[i] <- predictClass # save to the predictAs
}
```

## Step 5: Confusion Matrix

```{r, warning=FALSE}
confusionMatrix <- table(results$predictAs, results$Species) # get matrix
print(confusionMatrix) # print the confusion matrix
```

## Step 6: Calculate the accuracy rate and compare with homework 3

```{r, warning=FALSE}
accuracy = (confusionMatrix[1,1]+confusionMatrix[2,2]+confusionMatrix[3,3]) / 50
print("The accuracy is:")
print(accuracy)
```

Compare with my homework 3, the accuracy in homework 4 is higher.
