---
title: "Homework 2"
author: "Jiawen Qi (jiq10)"
date: "September 25, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Download the iris data from courseweb

I created a project for this homework, so the dataset can be read directly by the name without the path. When reading the csv file, `header = FALSE` because there is no column name in the original dataset. `stringAsFactor = FALSE` because I don't want the last column, Species, to be changed into factors automatically.

```{r}
iris <- read.csv("iris.csv", header = FALSE, stringsAsFactors = FALSE)
```

Using `dim()` to check the dimension of iris dataset.

```{r}
dim(iris)
```

From the result, we can know that there are 150 observations and 5 columns. So let's see the first several observations in this dataset using `head()`. And get a summary using `summary()`.

```{r}
head(iris)
summary(iris)
```

We need to rename the columns. Use `names()`.

```{r}
names(iris) <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")
```

## 2. Delete the setosa iris cases

First, let's check the species in this dataset using `table()`.

```{r}
table(iris$Species)
```

From the result, we can know that there are 50 Iris-setosa, 50 Iris-versicolor, and 50 Iris-virginica. Now we will delete the Iris-setosa cases by `subset()` and then save to a new data frame `iris_1`.

```{r}
iris_1 <- subset(iris, !(Species %in% c("Iris-setosa")))
```

Check the first several observations of `iris_1`.

```{r}
head(iris_1)
```

## 3. Keep petal width and sepal width and iris type.

To keep these three columns, we can use function, `subset()`. Then save to a new variable `iris_2`.

```{r}
iris_2 <- subset(iris_1, select = c(Sepal.Width, Petal.Width, Species))
```

Check the `iris_2` using `head()`.

```{r}
head(iris_2)
```

## 4. Make a dummy variable for iris type

To make a dummy variable, there is a package called `dummies` that can help me. Open that package first.

```{r}
library(dummies)
```

Then, create the dummy for iris type, Species. The function is `dummy()`. Save to the variable `dummyForSpecies`, and change to a data frame.

```{r}
dummyForSpecies <- dummy(iris_2$Species)
dummyForSpecies <- as.data.frame(dummyForSpecies)
```

Actually, we only need one column in dummyForSpecies because there are only to levels: Iris-versicolor and Iris-virginica. If it is not Iris-versicolor, it must be Iris-virginica. To make the dataset clear enough, I will keep both.Let's check the first several observations of `dummyForSpecies`.

```{r}
names(dummyForSpecies) <- c("SpeciesIris-versicolor", "SpeciesIris-virginica")
head(dummyForSpecies)
```

Combine the `iris_2` and `dummyForSpecies` using `cbind.data.frame()`. And save to a new varibale `iris_3`.

```{r}
iris_3 <- cbind.data.frame(iris_2, dummyForSpecies)
head(iris_3)
```

## 5. Create a random number, sort the data by that random number

Create a column to store random number, generate a random number for each observation and then sort by the random number column. 

```{r}
set.seed(66666)
iris_3$randomNumber <- 0
for (i in 1:length(iris_3$randomNumber)) {
  iris_3$randomNumber[i] <- runif(1)
}
iris_4 <- iris_3[order(iris_3$randomNumber),]
head(iris_4)
```

## 6.  Divide the data (100 obs) into 10 groups based on their order

We need to remove the randomNumber column using `subset()`, save as `iris_5`

```{r}
iris_5 <- subset(iris_4, select = -randomNumber)
head(iris_5)
```

Then, the index should be renamed from 1 to 100 using `rownames()`

```{r}
rownames(iris_5) <- c(1:length(iris_5$Species))
head(iris_5)
```

Divide into 10 groups.

```{r}
group_1 <- iris_5[1:10,]
group_2 <- iris_5[11:20,]
group_3 <- iris_5[21:30,]
group_4 <- iris_5[31:40,]
group_5 <- iris_5[41:50,]
group_6 <- iris_5[51:60,]
group_7 <- iris_5[61:70,]
group_8 <- iris_5[71:80,]
group_9 <- iris_5[81:90,]
group_10 <- iris_5[91:100,]
```

## 7. Take 9 of the ten groups and run a linear regression with petal width as a function of iris type and sepal width

Here, I choose group 1-9 and combine to a training data frame.

```{r}
training <- rbind.data.frame(group_1, group_2, group_3, group_4, group_5, group_6, group_7, group_8, group_9)
```

Run a linear regression model to predict petal width by iris type and sepal width. Actuallt, `lm()` in R will create dummy variable automatically. So we can use `Species` directly.

```{r}
model_1 <- lm(Petal.Width ~ Species + Sepal.Width, data = training)
summary(model_1)
```

From this model, we can conclude a predict equation that `Petal.Width = 0.13420 + 0.60056(when species is Iris-virginica) + 0.43203*Sepal.Width`

## 8. Do that 10 times so that each group of ten obs is omitted from one of the regressions

To make the linear regression 10 times, a for loop is a good idea to make it automatically.

```{r}
for (i in 1:10) {
  beginIndex <- i + 10 * (i - 1)
  endIndex <- beginIndex + 9
  testing <- iris_5[beginIndex:endIndex,]
  training <- iris_5[!(row.names(iris_5) %in% row.names(testing)),]
  model <- lm(Petal.Width ~ Species + Sepal.Width, data = training)
  cat("model", i, "Coefficients:", "\n")
  cat("Intercept:", model$coefficients[1],"\n")
  cat("SpeciesIris-virginica:", model$coefficients[2],"\n")
  cat("Sepal.Width :", model$coefficients[3],"\n")
  cat("\n")
}
```

## 9. Keep track of the intercept, dummy variable coefficient, and sepal width coefficient

To record the intercept, dummy variable coefficient, and sepal width coefficient, we need to create to three variables to save the value. Then, do the loop.

```{r}
Intercept <- numeric(10)
SpeciesIris_virginica <- numeric(10)
Sepal.Width <- numeric(10)
for (i in 1:10) {
  beginIndex <- i + 10 * (i - 1)
  endIndex <- beginIndex + 9
  testing <- iris_5[beginIndex:endIndex,]
  training <- iris_5[!(row.names(iris_5) %in% row.names(testing)),]
  model <- lm(Petal.Width ~ Species + Sepal.Width, data = training)
  Intercept[i] = model$coefficients[1]
  SpeciesIris_virginica[i] <- model$coefficients[2]
  Sepal.Width[i] <- model$coefficients[3]
}
cat("The Intercept values:", "\n")
cat(Intercept, "\n")
cat("\n")
cat("The SpeciesIris-virginica values:", "\n")
cat(SpeciesIris_virginica, "\n")
cat("\n")
cat("The Sepal.Width values:", "\n")
cat(Sepal.Width, "\n")
```

## 10. Make a histogram (hist) of those three values (there will be 3 histograms with ten observations in each histogram).

To draw histograms, `hist()` is a good function to use.

```{r}
hist(Intercept)
hist(SpeciesIris_virginica)
hist(Sepal.Width)
```
